New lower bounds for Schur and weak Schur numbers

### 2021-09-22 article/weakschur.tex
fix refs + precisions

\qquad In this section, we generalize Rowley's constructions for weak Schur numbers \cite{RowleyWS} and give an analogous
for weak Schur numbers of Abbott and Hanson's construction for Schur numbers. By analogy with the previous section,
we then introduce WS-templates, standing for \textit{"weak Schur templates"}, as well as an associated sequence \(\WS^+(n)\). 
We find templates and use them to provide new lower bounds for weak Schur numbers. Finally, we give a short explanation for 
the new lower bound \(\WS(6) \geqslant 646\) which was not directly obtained with a template contrary to the other lower bounds 
given in this paper.

@@

\qquad Up to now, no equivalent for weak Schur numbers of Abbott and Hanson's construction for Schur numbers
\cite{AbbottHanson} was known. Here we give a general lower bound for weak Schur numbers as a function of both
regular and weak Schur numbers. The following theorem, inspired by Rowley's inequalities for \(\WS (n+1)\) and
\(\WS (n+2)\), was found and proved by Romain Ageron.

\begin{theorem}
\label{Theorem3.1}
\begin{sloppypar}
Let \((p,k), (q,n) \in (\mathbb{N}^*)^2\). If there exists a partition of \([\![1,q]\!]\) into \(n\) weakly sum-free
subsets and a partition of \([\![1,p]\!]\) into \(k\) sum-free
subsets then there exists a partition of \({[\![1,p(q+\left \lceil \frac{q}{2} \right \rceil + 1)+q]\!]}\) into \(n+k\)
weakly sum-free subsets.
\end{sloppypar}
\end{theorem}

In particular, by setting \(q = \WS (n)\) and \(p = S(k)\) in \hyperref[Theorem3.1]{Theorem 3.1}, one obtains the following corollary.

\begin{corollary}
\label{Corollary3.2}
\( \forall (n,k) \in (\mathbb{N}^*)^2 \text{, } \WS (n+k) \geqslant S(k) \left (\WS (n) + \left \lceil \displaystyle \frac{\WS (n)}{2}
\right \rceil +1 \right) + \WS (n)\)
\end{corollary}

@@

\begin{remark}
The inequality from \hyperref[Corollary3.2]{Corollary 3.2} can be improved by adding 1 to the lower bound if \(\WS (n)\) is odd (more generally if \(q\) is
odd in the theorem). However, it lengthens the proof and it is never useful in practice.
\end{remark}

Given that \hyperref[Theorem3.1]{Theorem 3.1} will appear as a particular case of a more general theorem after the introduction of
templates for weak Schur numbers, we only give here an intuitive explanation of the demonstration; a formal
\hyperref[PreuveThm]{proof} using templates for weak Schur numbers is provided.

@@

\begin{proof}[\textsc{Proof.}]
\(\pi(y) \geqslant b + 1\) and thus \(b + 1 \leqslant x + \pi(y) \leqslant a + b\).
\[
\begin{array}{l l l l}
	\pi(x + y) & = & \pi(x + \pi(y)) & \text{by \hyperref[prop2]{Proposition 3.6}} \\
	 & = & x + \pi(y) & \text{by \hyperref[prop1]{Proposition 3.5}}
\end{array}
\]
\end{proof}

@@

Function \(\lambda\) maps an element \(x\) to its line number as mentioned in the intuitive explanation.
As we just did with \(\pi\), we proove three more results on both \(\pi\) and \(\lambda\) that will appear useful in
\hyperref[ConstructionWS]{Subsection 3.3}.

@@

\begin{proof}[\textsc{Proof.}]
By applying \hyperref[prop5]{Proposition 3.10} twice, we get \(a \lambda(x + y) + \pi(x + y) - a = x + y = x + a \lambda(y) + \pi(y) - a\).
The result is then obtained by simplying the equality. \\
\end{proof}

@@

\begin{proof}[\textsc{Proof.}]
By \hyperref[prop5]{Proposition 3.10}, \(x + y = a (\lambda(x) + \lambda(y)) + \pi(x) + \pi(y) - 2 a\). Then
\[
\begin{array}{l l l}
	\lambda(x + y) & = & \left \lfloor \dfrac{x + y - b - 1}{a} \right \rfloor + 1 \vspace{1mm} \\
	 & = & \left \lfloor \dfrac{a (\lambda(x) + \lambda(y)) + \pi(x) + \pi(y) - 2 a - b - 1}{a} \right \rfloor + 1 \vspace{1mm} \\
	 & = & \lambda(x) + \lambda(y) - 1 \left \lfloor \dfrac{\pi(x) + \pi(y) - b - 1}{a}\right\rfloor \\
	 & = & \lambda(x)+\lambda(y) -1 +1 ~~~~\text{since } \pi(x) + \pi(y) \in [\![a + b + 1, 2 a + b]\!] \\
	 & =& \lambda(x)+\lambda(y)
\end{array}
\]
\end{proof}

@@

\begin{proof}[\textsc{Proof.}]
The lower bound comes from the analogous of Abott and Hanson's construction for weak Schur numbers.
The upper bound comes from the fact that a WS-template with width \(a\) and \(n\) colors contains a partition of
\([\![1, a]\!]\) into \(n\) sum-free subsets. \\
\end{proof}

@@

Theorem provides a way to extend a S-template into a larger Schur partitions using other Schur partitions. Likewise, a WS-template can 
be extended into larger weak Schur partitions in using Schur partitions. This is the object of the following theorem.

\begin{theorem}
\label{Theorem3.18}
Let \((a,n,b) \in (\mathbb{N}^*)^3\) with \(a > b\) and \( (p,k) \in (\mathbb{N}^*)^2\). If there exists a partition of \([\![1,p]\!]\)
into \(k\) sum-free subsets and a \(b\)-WS-template \((A_1,...,A_{n+1})\) with width \(a\) and \(n+1\) colors,
then there exists a partition of \([\![1, p a + b]\!]\) into \(k+n\) weakly sum-free subsets.
\end{theorem}

In particular, by setting \(p = S(k)\) and \(a = \WS^+(n + 1)\) in \hyperref[Theorem3.18]{Theorem 3.18}, the next corollary follows.

\begin{corollary}
\label{Corollary3.19}
Let \(n,k \in \mathbb{N}^*\) and set \( b_{max} = \max \{b\in \mathbb{N}^*/ \WS_b^+(n+1) = \WS^+(n+1)\}\).
Then :
\[
\WS(n+k) \geqslant S(k) \WS^+(n+1)+b_{max}.
\]
\end{corollary}

@@

Split \([\![1, p a + b]\!]\) into three subsets. \\
NB: To keep the notation short, the restriction to \([\![b + 1, p a + b]\!]\) of \(\pi\) defined in \hyperref[DefinitionWS+]{Subsection 3.2}
is denoted by \(pi\) in the hereunder equations.

@@

\noindent \underline{\textbf{Case 1:} \((x,y) \in \mathcal{T}^2\)}
\par
If \(x + y \leqslant b\) then \(h(x+y)=f(x+y)\). Otherwise, \(b < x+y < a+b\) since \(b < a\) and
therefore \(\pi(x + y) = x +y\) (\hyperref[prop1]{Proposition 3.5}). Hence in both cases \(h(x+y)=f(x+y)\). Given that
\(f\) is a weakly sum-free coloring, \(f(x + y) \neq f(x)\) since \(f(x)=h(x)=h(y)=f(y)\) and \(x \neq y \). That
is \(h(x + y) \neq h(x)\). \\

\noindent \underline{\textbf{Case 2:} \((x,y) \in \mathcal{T} \times \mathcal{C}\)}
\par
Given that \(h(x) = h(y)\) and by definition of \(h\), \(f(x) = f(\pi(y))\). Besides, \(f(\pi(y)) \leqslant n\) since
\(y \in \mathcal{C}\). Two cases are to be distinguished according to the value of \(x + \pi(y)\).
\begin{itemize}
\item \begin{sloppypar}
	If \(x + \pi(y) \leqslant a + b\) then \(f(x + \pi(y)) = f(\pi(x + y))\) (\hyperref[prop3]{Proposition 3.7}). Given that \(f\) is
	a weakly sum-free coloring, \(f(x+\pi(y)) \neq f(x)\) since \(f(x)=f(\pi(y))\) and \(x \neq \pi(y)\) since
	\({x \leqslant b < \pi(y)}\).
	\end{sloppypar}
\item \begin{sloppypar}
	If \(x+\pi(y)> a+b\) then given that \(f\) is a WS-template and since \({f(x) = f(\pi(y)) \leqslant n}\),
	\({f(\pi(x+\pi(y))) \neq f(x)}\).~Furthermore \({f(\pi(x+\pi(y))) = f(\pi(x+y))}\) (\hyperref[prop2]{Proposition 3.6}), such that
	\({f(\pi(x+ y)) \neq f(x)}\).
	\end{sloppypar}
\end{itemize}
\par
Hence in both cases \(f(\pi(x+y)) \neq f(x)\). If  \(f(\pi(x+y)) \leqslant n\) then \(h(x+y) = f(\pi(x+y))\). Therefore
\(h(x+y) \neq h(x)\) since \(f(x) = h(x)\). Otherwise, \(f(\pi(x+y)) = n + 1\) and thus \(h(x+y) > n\). In particular,
\(h(x + y) \neq h(x)\) since \(h(x) = h(y) \leqslant n\). \\

\noindent \underline{\textbf{Case 3:} \((x,y) \in \mathcal{T} \times \mathcal{R}\)}
\par
Necessarily \(h(x) = h(y) = n + 1\). Two cases are to be distinguished according to the value of \(\lambda(x+y)\).
\begin{itemize}
\item If \(\lambda(y)=\lambda(x+y)\) then \(\pi(x + y) = x + \pi(y)\) (\hyperref[prop6]{Proposition 3.11}). By definition of
	\(h\), \(f(x) = f(\pi(y))\). Given that \(f\) is a weakly sum-free coloring, \(f(x + \pi(y)) \neq f(x)\) since
	\(f(x) = f(\pi(y))\) and \(x \neq \pi(y)\) since \({x \leqslant b < \pi(y)}\). Hence \(h(x + y) \neq h(x)\).
\item If \(\lambda(y) \neq \lambda(x + y)\) then \(\lambda(x + y) = \lambda(y) + 1\) since \(x \leqslant b < a\).
	Besides, \(n + 1 = h(y) = n +  g(\lambda(y))\). Hence \(g(\lambda(y)) = 1\). Moreover \(g(1) = 1\) since \(g\)
	is an ordered coloring. Therefore, given that \(g\) is sum-free, \(g(\lambda(y) + 1) \neq 1\). If \(\pi(x + y) \in
	A_{n + 1}\) then \(h(x + y) = n + g(\lambda(x + y)) \neq n + 1\). Otherwise, \(h(x + y) \leqslant n\). Hence
	in both cases \(h(x + y) \neq h(x)\).
\end{itemize}

\noindent \underline{\textbf{Case 4:} \((x,y) \in \mathcal{C}^2\)}
\par
By definition of \(h\) and since \(h(x)=h(y)\), \(f(\pi(x)) = f(\pi(y))\). Two cases are to be distinguished according
to the value of \(\pi(x)+\pi(y)\).
\begin{itemize}
\item If \(\pi(x) + \pi(y) \leqslant a+b\) then \(\pi(x)+\pi(y) = \pi(x + y)\). Hence \(f(\pi(x + y)) \neq f(\pi(x))\) since
	\(f\) is sum-free for \(x>b\)
\item \begin{sloppypar}
	If \(\pi(x)+\pi(y)>a+b\) then given that \(f\) is a WS-template, \({f(\pi(\pi(x)+\pi(y))) \neq f(\pi(x))}\) since
	\({f(\pi(x)) = f(\pi(y))}\). Besides,  \({f(\pi(\pi(x)+\pi(y))) = f(\pi(x + y))}\) (\hyperref[prop4]{Proposition 3.8}). Hence \({f(\pi(x + y))
	\neq  f(\pi(x))}\).
	\end{sloppypar}
\end{itemize}
\par
Hence in both cases \(f(\pi(x+y)) \neq f(x)\). If  \(f(\pi(x+y)) \leqslant n\) then \(h(x+y) = f(\pi(x+y))\). Therefore
\(h(x+y) \neq h(x)\) since \(f(x) = h(x)\). Otherwise, \(f(\pi(x+y)) = n + 1\) and thus \(h(x+y) > n\). In particular,
\(h(x + y) \neq h(x)\) since \(h(x) = h(y) \leqslant n\). \\

@@

\noindent \underline{\textbf{Case 6:} \((x,y) \in \mathcal{R}^2\)}
\par
In particular \(f(\pi(x)) = f(\pi(y))=n + 1\). Three cases are to be distinguished according to the value of \(\pi(x) + \pi(y)\).
\begin{itemize}
\item If \(\pi(x) + \pi(y) \in [\![a + b + 1, 2 a + b]\!]\) then \(\lambda(x + y) = \lambda(x) + \lambda(y)\)
	(\hyperref[prop7]{Proposition 3.12}). By definition of \(h\) and since \(h(x) = h(y)\), \(g(\lambda(x)) = g(\lambda(y))\).
	Hence, \(h(\lambda(x + y)) \neq h(\lambda(x))\) since \(h\) is a sum-free coloring. If \(f(x+y) \geqslant n + 1\)
	then \(h(x + y) = n + g(\lambda(x + y))\). And \(h(x) = n + g(\lambda(x))\). Therefore, \(h(x + y)  \neq h(x)\).
	Otherwise \(h(x+y) \leqslant n < h(x)\). In particular \(h(x + y) \neq h(x)\).
\item If \(\pi(x)+\pi(y)>2a+b\) then \(f(\pi(\pi(x)+\pi(y))) \neq f(\pi(x)) = n + 1\) since \(f\) is a \(b\)-WS template and
	\(f(\pi(x)) = f(\pi(y))\). Given that \(\pi(\pi(x)+\pi(y)) = \pi(x+y)\) (\hyperref[prop4]{Proposition 3.8}), \(f(\pi(x+y)) \neq n + 1\).
\item \begin{sloppypar}
	If \(\pi(x)+\pi(y)\leqslant b+a\) then, given that \(\pi(x)+\pi(y) \geqslant b\) and \(f_{| [\![b, a + b ]\!]}\) is
	sum-free, \({f(\pi(x) + \pi(y)) \neq f(\pi(x)) = n + 1}\). That is \({f(\pi(x + y)) \neq n + 1}\) (\hyperref[prop1]{Proposition 3.5}).
	\end{sloppypar}
\end{itemize}
\par
In both of the last two cases, \({f(\pi(x + y)) \neq n + 1}\) that is \(x+y \in \mathcal{C}\). Therefore \(h(x+y) < n \leqslant h(x)\).
In particular, \(h(x + y) \neq h(x)\). \\
\end{proof}

@@

\begin{theorem}
\label{Theorem3.21}
Let \((k,p) \in (\mathbb{N}^*)^2\) and \((a, n, b) \in (\mathbb{N}^*)^3\). If there exists a S-template with width
\(p\) and \(k+1\) colors and a \(b\)-WS-template with width \(a\) and \(n\) colors, then there exists a \(pb\)-WS-template
with width \(pq\) and \((n+k)\) colors.
\end{theorem}

\hyperref[Theorem3.21]{Theorem 3.21} yields the following corollary.

@@

\begin{proof}[\textsc{Proof.}]
The idea is the same as in the previous theorem. The only difference is the WS property inherited
from the WS-template. \\
\end{proof}

\hyperref[Theorem3.1]{Theorem 3.1} can be seen as a particular case of WS-template in the same way Abott and Hanson's 
construction can be seen as a particular case of S-template.

\begin{proof}[\textsc{Proof of \hyperref[Theorem3.1]{Theorem 3.1}.}]
\label{PreuveThm}
Let \((q, n) \in (\mathbb{N}^*)^2\) such that there exists a partition of \([\![1, q]\!]\) into n weakly sum-free subsets. 
Let \(f: [\![1, q]\!] \rightarrow [\!]1, n]\!]\) a weakly sum-free colouring.
Let \(b=q\) and \(a = q + \left \lceil \dfrac{q}{2} \right \rceil + 1\). A new colouring \(g\) is defined as follows:

@@

The result is then obtained by applying \hyperref[Theorem3.18]{Theorem 3.18}. \\
\end{proof}

As in \hyperref[Corollary2.5]{Corollary 2.5}, the additive constant \hyperref[Theorem3.18]{Theorem 3.18} can be improved 
by weakening the hypotheses made on the last row. The principle behind it is the same as in \hyperref[Proposition2.6]{Proposition 2.6}.

@@

We also have a similar theorem where only \(\WS^+\) is involved.

@@

\par
Here are the inequalities given by the current best WS-templates (to our knowledge). The \hyperref[WS-templates]{templates} 
corresponding to the third inequality can be found in the appendix.

@@

Inequalities (\ref{WS(n+1)}) and (\ref{WS(n+2)}) were found by Rowley, they are detailed in \cite{RowleyWS}. 
Inequality (\ref{WS(n+3)}) cannot be further improved and was found with a SAT solver. It uses the first 
sophistication explained in \hyperref[ConstructionWS]{Subsection 3.3} in order to add
the last number in the first color. As for inequality (\ref{WS(n+4)}), it was obtained by combining a S-template with 
width 33 with a WS-template with width 4. The best template we could find with a computer search gives the 
inequality \(\WS(n+4) \geqslant 127 S(n) + 68\). It was also found with the SAT solver. In order to reduce the search space, 
we only looked for WS-templates of five colors which start with a near-optimal \(\WS(4)\) partition and we assumed that the 
special color was the last by order of appearance. We highly suspect that better WS-templates with \(n \geqslant 5\) 
colors can be found but one would have not to use the above assumptions. One may try to go over a different search space 
using a Monte-Carlo method, as in \cite{Bouzy2015AnAP}. This could be the suject of a future work.

Like in \hyperref[Subsection2.3]{Subsection 2.3}, we compute the lower bounds given by inequalities (\ref{WS(n+1)}),
(\ref{WS(n+2)}) and (\ref{WS(n+3)}) for \( n \in [\![8,15]\!] \). The best lower bound for each value of \(n\) is highlighted.

@@

\qquad We started by giving a new construction which can be seen as an equivalent for weakly sum-free partitions of Abott
and Hanson's construction for sum-free partitions. We then generalized this construction by introducing WS-templates. This
allows us to find new lower bounds and new inequalities for weak Schur numbers. One may notice the significant gap
between the former lower bounds for weak Schur numbers obtained by conducting a computer search and the new lower bounds
obtained with WS-templates.

>>>>

\qquad In this section, we generalize Rowley's constructions for weak Schur numbers \cite{RowleyWS} and give an analogous
for weak Schur numbers of Abbott and Hanson's construction for Schur numbers \cite{AbbottHanson}. By analogy with the previous section,
we then introduce WS-templates, standing for \textit{"weak Schur templates"}, as well as an associated sequence \(\WS^+(n)\). 
We find templates and use them to provide new lower bounds for weak Schur numbers. Finally, we give a short explanation for 
the new lower bound \(\WS(6) \geqslant 646\) which was not directly obtained with a template contrary to the other lower bounds 
given in this paper.

@@

\qquad Up to now, no equivalent for weak Schur numbers of Abbott and Hanson's construction for Schur numbers
\cite{AbbottHanson} was known. Here we give a general lower bound for weak Schur numbers as a function of both
regular and weak Schur numbers.

\begin{theorem}
\label{thm:WS-ab}
\begin{sloppypar}
Let \((p,k), (q,n) \in (\mathbb{N}^*)^2\). If there exists a partition of \([\![1,q]\!]\) into \(n\) weakly sum-free
subsets and a partition of \([\![1,p]\!]\) into \(k\) sum-free
subsets then there exists a partition of \({[\![1,p(q+\left \lceil \frac{q}{2} \right \rceil + 1)+q]\!]}\) into \(n+k\)
weakly sum-free subsets.
\end{sloppypar}
\end{theorem}

In particular, by setting \(q = \WS (n)\) and \(p = S(k)\) in Theorem \ref{thm:WS-ab}, one obtains the following corollary.

\begin{corollary}
\label{cor:ineqWS-S}
\( \forall (n,k) \in (\mathbb{N}^*)^2 \text{, } \WS (n+k) \geqslant S(k) \left (\WS (n) + \left \lceil \displaystyle \frac{\WS (n)}{2}
\right \rceil +1 \right) + \WS (n)\)
\end{corollary}

@@

\begin{remark}
The inequality from Corollary \ref{cor:ineqWS-S} can be improved by adding 1 to the lower bound if \(\WS (n)\) is odd (more generally if \(q\) is
odd in Theorem \ref{thm:WS-ab}). However, it lengthens the proof and it is never useful in practice.
\end{remark}

Given that Theorem \ref{thm:WS-ab} will appear as a particular case of a more general theorem after the introduction of
templates for weak Schur numbers, we only give here an intuitive explanation of the demonstration; a formal
\hyperref[PreuveThm]{proof} using templates for weak Schur numbers is provided in Subsection \ref{ConstructionWS}.

@@

\begin{proof}[\textsc{Proof.}]
\(\pi(y) \geqslant b + 1\) and thus \(b + 1 \leqslant x + \pi(y) \leqslant a + b\).
\[
\begin{array}{l l l l}
	\pi(x + y) & = & \pi(x + \pi(y)) & \text{by Proposition \ref{prop2}} \\
	 & = & x + \pi(y) & \text{by Proposition \ref{prop1}}
\end{array}
\]
\end{proof}

@@

Function \(\lambda\) maps an element \(x\) to its line number as mentioned in the intuitive explanation.
As we just did with \(\pi\), we prove three more calculative results on both \(\pi\) and \(\lambda\) that are used in
Subsection \ref{ConstructionWS}.

@@

\begin{proof}[\textsc{Proof.}]
By applying Proposition \ref{prop5} twice, we get \(a \lambda(x + y) + \pi(x + y) - a = x + y = x + a \lambda(y) + \pi(y) - a\).
The result is then obtained by simplying the equality. \\
\end{proof}

@@

\begin{proof}[\textsc{Proof.}]
By Proposition \ref{prop5}, \(x + y = a (\lambda(x) + \lambda(y)) + \pi(x) + \pi(y) - 2 a\). Then
\[
\begin{array}{l l l}
	\lambda(x + y) & = & \left \lfloor \dfrac{x + y - b - 1}{a} \right \rfloor + 1 \vspace{1mm} \\
	 & = & \left \lfloor \dfrac{a (\lambda(x) + \lambda(y)) + \pi(x) + \pi(y) - 2 a - b - 1}{a} \right \rfloor + 1 \vspace{1mm} \\
	 & = & \lambda(x) + \lambda(y) - 1 \left \lfloor \dfrac{\pi(x) + \pi(y) - b - 1}{a}\right\rfloor \\
	 & = & \lambda(x)+\lambda(y) -1 +1 ~~~~\text{since } \pi(x) + \pi(y) \in [\![a + b + 1, 2 a + b]\!] \\
	 & =& \lambda(x)+\lambda(y)
\end{array}
\]
\end{proof}

@@

\begin{proof}[\textsc{Proof.}]
The lower bound comes from Corollary \ref{cor:ineqWS-S}.
The upper bound comes from the fact that a WS-template with width \(a\) and \(n\) colors is in particular a partition of
\([\![1, a]\!]\) into \(n\) weakly sum-free subsets. \\
\end{proof}

@@

Theorem \ref{thm:Stemp} provides a way to extend a S-template into a larger Schur partitions using other Schur partitions. Likewise, a WS-template can 
be extended into larger weak Schur partitions by using Schur partitions. This is the object of the following theorem.

\begin{theorem}
\label{thm:WStemp}
Let \((a,n,b) \in (\mathbb{N}^*)^3\) with \(a > b\) and \( (p,k) \in (\mathbb{N}^*)^2\). If there exists a partition of \([\![1,p]\!]\)
into \(k\) sum-free subsets and a \(b\)-WS-template \((A_1,...,A_{n+1})\) with width \(a\) and \(n+1\) colors,
then there exists a partition of \([\![1, p a + b]\!]\) into \(k+n\) weakly sum-free subsets.
\end{theorem}

In particular, by setting \(p = S(k)\) and \(a = \WS^+(n + 1)\) in Theorem \ref{thm:WStemp}, the next corollary follows.

\begin{corollary}
\label{cor:ineqWS+}
Let \(n,k \in \mathbb{N}^*\) and set \( b_{max} = \max \{b\in \mathbb{N}^*/ \WS_b^+(n+1) = \WS^+(n+1)\}\).
Then :
\[
\WS(n+k) \geqslant S(k) \WS^+(n+1)+b_{max}.
\]
\end{corollary}

@@

Split \([\![1, p a + b]\!]\) into three subsets. \\
NB: To keep the notation short, the restriction to \([\![b + 1, p a + b]\!]\) of \(\pi\) defined in Subsection \ref{DefinitionWS+}
is denoted by \(\pi\) in the hereunder equations.

@@

\noindent \underline{\textbf{Case 1:} \((x,y) \in \mathcal{T}^2\)}
\par
If \(x + y \leqslant b\) then \(h(x+y)=f(x+y)\). Otherwise, \(b < x+y < a+b\) since \(b < a\) and
therefore \(\pi(x + y) = x +y\) (Proposition \ref{prop1}). Hence in both cases \(h(x+y)=f(x+y)\). Given that
\(f\) is a weakly sum-free coloring, \(f(x + y) \neq f(x)\) since \(f(x)=h(x)=h(y)=f(y)\) and \(x \neq y \). That
is \(h(x + y) \neq h(x)\). \\

\noindent \underline{\textbf{Case 2:} \((x,y) \in \mathcal{T} \times \mathcal{C}\)}
\par
Given that \(h(x) = h(y)\) and by definition of \(h\), \(f(x) = f(\pi(y))\). Besides, \(f(\pi(y)) \leqslant n\) since
\(y \in \mathcal{C}\). Two cases are to be distinguished according to the value of \(x + \pi(y)\).
\begin{itemize}
\item \begin{sloppypar}
	If \(x + \pi(y) \leqslant a + b\) then \(f(x + \pi(y)) = f(\pi(x + y))\) (Proposition \ref{prop3}). Given that \(f\) is
	a weakly sum-free coloring, \(f(x+\pi(y)) \neq f(x)\) since \(f(x)=f(\pi(y))\) and \(x \neq \pi(y)\) since
	\({x \leqslant b < \pi(y)}\).
	\end{sloppypar}
\item \begin{sloppypar}
	If \(x+\pi(y)> a+b\) then given that \(f\) is a WS-template and since \({f(x) = f(\pi(y)) \leqslant n}\),
	\({f(\pi(x+\pi(y))) \neq f(x)}\).~Furthermore \({f(\pi(x+\pi(y))) = f(\pi(x+y))}\) (Proposition \ref{prop2}), such that
	\({f(\pi(x+ y)) \neq f(x)}\).
	\end{sloppypar}
\end{itemize}
\par
Hence in both cases \(f(\pi(x+y)) \neq f(x)\). If  \(f(\pi(x+y)) \leqslant n\) then \(h(x+y) = f(\pi(x+y))\). Therefore
\(h(x+y) \neq h(x)\) since \(f(x) = h(x)\). Otherwise, \(f(\pi(x+y)) = n + 1\) and thus \(h(x+y) > n\). In particular,
\(h(x + y) \neq h(x)\) since \(h(x) = h(y) \leqslant n\). \\

\noindent \underline{\textbf{Case 3:} \((x,y) \in \mathcal{T} \times \mathcal{R}\)}
\par
Necessarily \(h(x) = h(y) = n + 1\). Two cases are to be distinguished according to the value of \(\lambda(x+y)\).
\begin{itemize}
\item If \(\lambda(y)=\lambda(x+y)\) then \(\pi(x + y) = x + \pi(y)\) (Proposition \ref{prop6}). By definition of
	\(h\), \(f(x) = f(\pi(y))\). Given that \(f\) is a weakly sum-free coloring, \(f(x + \pi(y)) \neq f(x)\) since
	\(f(x) = f(\pi(y))\) and \(x \neq \pi(y)\) since \({x \leqslant b < \pi(y)}\). Hence \(h(x + y) \neq h(x)\).
\item If \(\lambda(y) \neq \lambda(x + y)\) then \(\lambda(x + y) = \lambda(y) + 1\) since \(x \leqslant b < a\).
	Besides, \(n + 1 = h(y) = n +  g(\lambda(y))\). Hence \(g(\lambda(y)) = 1\). Moreover \(g(1) = 1\) since \(g\)
	is an ordered coloring. Therefore, given that \(g\) is sum-free, \(g(\lambda(y) + 1) \neq 1\). If \(\pi(x + y) \in
	A_{n + 1}\) then \(h(x + y) = n + g(\lambda(x + y)) \neq n + 1\). Otherwise, \(h(x + y) \leqslant n\). Hence
	in both cases \(h(x + y) \neq h(x)\).
\end{itemize}

\noindent \underline{\textbf{Case 4:} \((x,y) \in \mathcal{C}^2\)}
\par
By definition of \(h\) and since \(h(x)=h(y)\), \(f(\pi(x)) = f(\pi(y))\). Two cases are to be distinguished according
to the value of \(\pi(x)+\pi(y)\).
\begin{itemize}
\item If \(\pi(x) + \pi(y) \leqslant a+b\) then \(\pi(x)+\pi(y) = \pi(x + y)\). Hence \(f(\pi(x + y)) \neq f(\pi(x))\) since
	\(f\) is sum-free for \(x>b\)
\item \begin{sloppypar}
	If \(\pi(x)+\pi(y)>a+b\) then given that \(f\) is a WS-template, \({f(\pi(\pi(x)+\pi(y))) \neq f(\pi(x))}\) since
	\({f(\pi(x)) = f(\pi(y))}\). Besides,  \({f(\pi(\pi(x)+\pi(y))) = f(\pi(x + y))}\) (Proposition \ref{prop4}). Hence \({f(\pi(x + y))
	\neq  f(\pi(x))}\).
	\end{sloppypar}
\end{itemize}
\par
Hence in both cases \(f(\pi(x+y)) \neq f(x)\). If  \(f(\pi(x+y)) \leqslant n\) then \(h(x+y) = f(\pi(x+y))\). Therefore
\(h(x+y) \neq h(x)\) since \(f(x) = h(x)\). Otherwise, \(f(\pi(x+y)) = n + 1\) and thus \(h(x+y) > n\). In particular,
\(h(x + y) \neq h(x)\) since \(h(x) = h(y) \leqslant n\). \\

@@

\noindent \underline{\textbf{Case 6:} \((x,y) \in \mathcal{R}^2\)}
\par
In particular \(f(\pi(x)) = f(\pi(y))=n + 1\). Three cases are to be distinguished according to the value of \(\pi(x) + \pi(y)\).
\begin{itemize}
\item If \(\pi(x) + \pi(y) \in [\![a + b + 1, 2 a + b]\!]\) then \(\lambda(x + y) = \lambda(x) + \lambda(y)\)
	(Proposition \ref{prop7}). By definition of \(h\) and since \(h(x) = h(y)\), \(g(\lambda(x)) = g(\lambda(y))\).
	Hence, \(h(\lambda(x + y)) \neq h(\lambda(x))\) since \(h\) is a sum-free coloring. If \(f(x+y) \geqslant n + 1\)
	then \(h(x + y) = n + g(\lambda(x + y))\). And \(h(x) = n + g(\lambda(x))\). Therefore, \(h(x + y)  \neq h(x)\).
	Otherwise \(h(x+y) \leqslant n < h(x)\). In particular \(h(x + y) \neq h(x)\).
\item If \(\pi(x)+\pi(y)>2a+b\) then \(f(\pi(\pi(x)+\pi(y))) \neq f(\pi(x)) = n + 1\) since \(f\) is a \(b\)-WS template and
	\(f(\pi(x)) = f(\pi(y))\). Given that \(\pi(\pi(x)+\pi(y)) = \pi(x+y)\) (Proposition \ref{prop4}), \(f(\pi(x+y)) \neq n + 1\).
\item \begin{sloppypar}
	If \(\pi(x)+\pi(y)\leqslant b+a\) then, given that \(\pi(x)+\pi(y) \geqslant b\) and \(f_{| [\![b, a + b ]\!]}\) is
	sum-free, \({f(\pi(x) + \pi(y)) \neq f(\pi(x)) = n + 1}\). That is \({f(\pi(x + y)) \neq n + 1}\) (Proposition \ref{prop1}).
	\end{sloppypar}
\end{itemize}
\par
In both of the last two cases, \({f(\pi(x + y)) \neq n + 1}\) that is \(x+y \in \mathcal{C}\). Therefore \(h(x+y) < n \leqslant h(x)\).
In particular, \(h(x + y) \neq h(x)\). \\
\end{proof}

@@

\begin{theorem}
\label{thm:compWStemp}
Let \((k,p) \in (\mathbb{N}^*)^2\) and \((a, n, b) \in (\mathbb{N}^*)^3\). If there exists a S-template with width
\(p\) and \(k+1\) colors and a \(b\)-WS-template with width \(a\) and \(n\) colors, then there exists a \(pb\)-WS-template
with width \(pq\) and \((n+k)\) colors.
\end{theorem}

Theorem \ref{thm:compWStemp} yields the following corollary.

@@

\begin{proof}[\textsc{Proof.}]
The idea is the same as in the previous theorem. The only difference is the WS property inherited
from both the S-template and the WS-template. \\
\end{proof}

Theorem \ref{thm:WS-ab} can be seen as a particular case of WS-template in the same way Abott and Hanson's 
construction \cite{AbbottHanson} can be seen as a particular case of S-template.

\begin{proof}[\textsc{Proof of Theorem \ref{thm:WS-ab}.}]
\label{PreuveThm}
Let \((q, n) \in (\mathbb{N}^*)^2\) such that there exists a partition of \([\![1, q]\!]\) into n weakly sum-free subsets. 
Let \(f: [\![1, q]\!] \rightarrow [\!]1, n]\!]\) a weakly sum-free colouring.
Let \(b=q\) and \(a = q + \left \lceil \dfrac{q}{2} \right \rceil + 1\). A new colouring \(g\) is defined as follows:

@@

The result is then obtained by applying Theorem \ref{thm:WStemp}. \\
\end{proof}

As in Corollary \ref{cor:ineqS}, the additive constant Theorem \ref{thm:WStemp} can be improved 
by weakening the hypotheses made on the last row. The principle behind it is the same as in Proposition \ref{prop:rafStemp}.

@@

@@

\par
Here are the inequalities given by the current best WS-templates. The \hyperref[WS-templates]{templates} 
corresponding to the third inequality can be found in the appendix.

@@

Inequalities (\ref{WS(n+1)}) and (\ref{WS(n+2)}) were found by Rowley, they are detailed in \cite{RowleyWS}. 
Inequality (\ref{WS(n+3)}) cannot be further improved and was found with a SAT solver. It uses the first 
sophistication explained in Subsection \ref{ConstructionWS} in order to add
the last number in the first color. As for inequality (\ref{WS(n+4)}), it was obtained by combining a S-template with 
width 33 with a WS-template with width 4. The best template we could find with a computer search gives the 
inequality \(\WS(n+4) \geqslant 127 S(n) + 68\). It was also found with the SAT solver. In order to reduce the search space, 
we only looked for WS-templates of five colors which start with a near-optimal \(\WS(4)\) partition and we assumed that the 
special color was the last by order of appearance. We highly suspect that better WS-templates with \(n \geqslant 5\) 
colors can be found but one would have not to use the above assumptions. One may try to go over a different search space 
using a Monte-Carlo method, as in \cite{Bouzy2015AnAP}. This could be the suject of a future work.

Like in  Subsection \ref{subsec:lowS}, we compute the lower bounds given by inequalities (\ref{WS(n+1)}),
(\ref{WS(n+2)}) and (\ref{WS(n+3)}) for \( n \in [\![8,15]\!] \). The best lower bound for each value of \(n\) is highlighted.

@@

\qquad We started by giving a new construction which can be seen as an equivalent for weakly sum-free partitions of Abott
and Hanson's construction for sum-free partitions. We then generalized this construction by introducing WS-templates. This
allows us to find new lower bounds and new inequalities for weak Schur numbers. One may notice the significant gap
between the former lower bounds for weak Schur numbers obtained by conducting a computer search and the new lower bounds
obtained with WS-templates. We reckon better WS-templates with \(n \geqslant 5\) colors can be found by making different assumptions 
and using a different method (Monte-Carlo methods for instance).
